Comprehensive Quality-Aware Automated Semantic Web Service Composition

### 2017-05-11 paper.tex
update

%where PSO optimise a queue of task-related services that is decoded into composition solutions. We compare the effectiveness of PSO-based method with one recent GP-based method using the proposed comprehensive quality model, and also compare the effectiveness of our proposed comprehensive quality model with one promising QoS model.

@@

\textit{Web service composition} aims to loosely couple a set of web services to provide a value-added composite service that accommodates complex functional and non-functional requirements of service users. 
%This application is developed by integrating interoperable and collaborative functionalities over heterogeneous systems. 
Two most notable challenges for web service composition are ensuring interoperability of services and achieving Quality of Service (QoS) optimisation \cite{fensel2011semantic}. \textit{Interoperability} of web services presents challenge in syntactic and semantic dimensions. The syntactic dimension is covered by the XML-based technologies, such as $WSDL$, $SOAP$. The semantic dimension enables a better collaboration through ontology-based semantics, such as OWL-S, WSML, and SAWSDL \cite{petrie2016web}. \textit{Semantic web services composition} is distinguished from the syntactic service composition, the resources of semantic web services are described semantically to enable a better interoperability for chaining web services. Another challenge is related to QoS optimisation. This problem gives birth to \textit{QoS-aware service composition} that aims to find composition solutions with optimised QoS.

Existing works on service composition focus mainly on addressing only one challenge above. In these works, huge efforts have been devoted to QoS-aware web service compositions assuming a pre-defined abstract workflow is given. This is generally considered as a \textit{semi-automated web service composition} approach. Generating composition plans automatically in discovering and selecting suitable web services is a NP-hard problem \cite{moghaddam2014service}. In the past few years, many approaches \cite{gupta2015optimization,ma2015hybrid,qi2010combining,da2016particle,da2015graphevol,yu2013adaptive} to QoS-aware web service composition employ Evolutionary Computation (EC) techniques to automatically generate composition solutions. Genetic Programming (GP) based approaches produce promising results, but these approaches often require repairing or penalising the solutions \cite{ma2015hybrid,yu2013adaptive}. However, Particle Swarm Optimisation (PSO) is considered to be an easy way to maintain the correctness of solutions in solving combinatorial optimisation problems \cite{da2016particle}. All these works have enabled an \textit{automatic web service composition}, but do not optimise QoS and quality of semantic matchmaking simultaneously to achieve a desirable balance on both sides.

The overall goal of this paper is to \textit{develop a PSO-based approach to comprehensive quality-aware automated semantic web service composition that satisfactorily optimises both QoS and semantic matchmaking quality}. Particularly, this paper extends existing works on QoS-aware service composition by considering jointly optimising QoS and semantic matchmaking quality, which is proposed as a comprehensive quality model. We will achieve three objectives in this work:

@@

 \item To propose a PSO-based service composition approach using the proposed comprehensive quality model. To do that, we aim to find a service candidate queue that can be decoded into a service composition with near-optimal comprehensive quality.  
 \item To address the effectiveness of our PSO-based approach and a desirable balance achieved using our comprehensive quality model, we first compare our PSO-based approach with one recent GP-based approach \cite{ma2015hybrid} using our proposed quality model, and then compare our proposed quality model with one widely used QoS model using our proposed PSO-based approach.
\end{enumerate}
\vspace{-0.5cm}
\section{Related Work} \label{relatedWork}
While web service composition has attracted much research over the last decade, most efforts focus on either semantic web service composition \cite{bansal2016generalized,boustil2014semantic,mier2015integrated} or QoS-aware web service composition \cite{gupta2015optimization,ma2015hybrid,qi2010combining,da2016particle,da2015graphevol,yu2013adaptive}, Only very few works address both semantic matchmaking quality and QoS for web service composition problems. To the best of our knowledge, \cite{fanjiang2014semantic,lecue2009optimizing,pop2009immune} reported about first attempts that consider both aspects together.

@@

In summary, despite a large number of approaches for semantic web service composition and QoS-aware service composition approaches, there is a lack of a fully automated semantic service composition approach to optimise semantic matchmaking quality and QoS simultaneously. 
%Therefore, we first propose an automated comprehensive quality-aware semantic web service composition approach. To do that, we will design a comprehensive quality model that considers both QoS and quality of semantic matchmaking between services. We will then propose a PSO-based approach to this semantic service composition using our proposed comprehensive quality model to evaluate candidate solutions.  
\vspace{-0.3cm}

\section{Problem Description and Comprehensive Quality Model}\label{Motivation and Problem Description}

@@

Given a service request $T=(I_T,O_T)$, we represent a service composition solution for $T$ with services $S_1,\ldots,S_n$  by a weighted DAG, $WG=(V,E)$ with node set $V=\{Start, S_1, S_2, \ldots, S_n, End\}$ and edge set $E = \{e_{1}, e_{2},... e_{m} \}$. $Start$ and $End$ are two special services defined as $Start = (\emptyset, I_T, \emptyset )$ and $End  = (O_T, \emptyset, \emptyset)$ that account for the input and output requirements given by the request. Each edge $e$ from a service $S$ to a service $S'$ means that service $S$ produces an output $a\in O_S$ that is matched ($exact$ or $plugin$) to an input $b\in I_{S'}$ to be consumed by service $S'$ in the composition. Based on the matchmaking type, the \emph{semantic matchmaking quality} of edge $e$ can be defined as follows:
\vspace{-0.2cm}
\begin{equation}
\label{equation10}
type_e = 
\begin{cases}
	1 & \text{ if $a\equiv b$ ($exact$ match)},\\
	p & \text{ if $a \sqsubseteq b$ ($plugin$ match)}
\end{cases}
\end{equation}
\begin{equation}
\label{equation11}
sim_e = sim(a,b) = \frac{2N_c}{N_{a}+N_{b}}
\end{equation}

\noindent with a suitable parameter $p, 0<p< 1$, see Section \ref{experiment_study}. However, if more than one pair of matched output and input exist from service $S$ to service $S'$, $type_e$ and $sim_e$ will take on their average values.

The \emph{semantic matchmaking quality} of the service composition can be obtained by aggregating over all edges in $E$ as follows:
%\begin{equation}
%\label{eqSM}
%SM \stackrel{.}{=} (MT, SIM)
%\end{equation}
\vspace{-0.2cm}
\begin{equation}
\label{equation6}
MT {=} \prod_{j=1}^{m} type_ {e_{j}}
\end{equation}
\begin{equation}
\label{equation7}
SIM {=} \frac{1}{m}\sum_{j=1}^m sim_ {e_{j}}
\end{equation}
%\begin{equation}
%\label{eqCQ}
%CQ \stackrel{.}{=} (MT, \  SIM, \  A,\  R,\  C,\  T)
%\end{equation}
The \emph{QoS} of the service composition can be obtained by aggregating the QoS values of the participating services \cite{ma2015hybrid}. For a service composition with services $ S_1, S_2, ... S_n$ we obtain the reliability $R=\prod\limits^n_{k=1}r_{S_k}$, the availability $A=\prod\limits^n_{k=1}a_{S_k}$, the cost $C=\sum\limits^n_{k=1}c_{S_k}$, and the response time $T$ is the time of most time-consuming path in the composition, i.e., $$T = MAX \{\sum\limits^{\ell_j}_{k=1}t_{S_k} | \text{ $j\in\{1,\ldots,m\}$ and $\ell_j$ is the length of a path $P_j$}\}.$$

%Consequently, in Eq. \ref{eqCQ}, the comprehensive quality $CQ$ consists of the aggregated semantic matchmaking quality $SM$ and the QoS calculation in DAG discussed before.
\vspace{-0.1cm}
\noindent When multiple quality criteria are involved into decision making, then the overall fitness of a solution can be defined as a weighted sum of the individual criteria: 
\vspace{-0.2cm}
\begin{equation}
\label{equation8}
Fitness = w_1 \hat{MT} + w_2 \hat{SIM} + w_3 \hat{A} + w_4 \hat{R} + w_5(1 - \hat{T}) + w_6(1 - \hat{C})
\end{equation}
\noindent with $\sum_{k=1}^{6} w_k= 1$. We call this objective function the \emph{comprehensive quality model} for service composition.
The weights can be adjusted according to users' preferences. We normalise $MT$, $SIM$, $A$, $R$, $T$, and $C$ so that the function value falls within the range from 0 to 1 using Eq. (\ref{equation9}). To simplify the presentation we also use the notation $(Q_1,Q_2,Q_3,Q_4,Q_5,Q_6) = (MT,SIM,A,R,T,C)$. $MT$ and $SIM$ have minimum value 0 and maximum value 1. The minimum and maximum value of $A$, $R$, $T$, and $C$ are calculated across all task-related candidates in the service repository $\mathcal{SR}$ using the greedy search in \cite{ma2015hybrid,da2016genetic}.

@@

\section{PSO-based Approach to Comprehensive Quality-Aware Automated Semantic Web Service Composition}\label{qswsc_approach}
\vspace{-0.8cm}
\begin{figure}[h]
\centering
\fbox{\includegraphics[scale=.4]{overview.pdf}}
 \caption{An overview of our PSO-based approach to comprehensive quality-aware automated semantic web service composition.}
 \label{overview}
\end{figure}
\vspace{-1.0cm}
\subsection{An Overview of our PSO-based Approach}\label{PSO_based_approach}

As PSO has shown promise in solving combinatorial optimisation problems, we propose a PSO-based approach to comprehensive quality-aware automated semantic web service composition. Fig. \ref{overview} shows an overview of our approach consisting of four steps:

@@

Recently, a QoS Model, $Fitness = w_1 \hat{A} + w_2 \hat{R} + w_3(1 - \hat{T}) + w_4(1 - \hat{C})$, where $\sum_{i=1}^{4} w_i = 1$, is widely used for QoS-aware web service composition \cite{ma2015hybrid,da2016particle,da2015graphevol}. To show the effectiveness of our proposed comprehensive quality model, we compare the best solutions found by this QoS model and our comprehensive model using our PSO-based approach. We record and compare the mean values of both $SM$ ($SM = 0.5 \hat{MT} + 0.5 \hat{SIM}$) and $QoS$($QoS = 0.25 \hat{A} + 0.25 \hat{R} + 0.25(1 - \hat{T}) + 0.25(1 - \hat{C})$) of best solutions over 30 independent runs. To make the comparison informative, all these recorded values have been normalised from 0 to 1, and compared using independent-samples t tests, see Table \ref{decisionTable}.

>>>>

@@

\textit{Web service composition} aims to loosely couple a set of web services to provide a value-added composite service that accommodates complex functional and non-functional requirements of service users. 
%This application is developed by integrating interoperable and collaborative functionalities over heterogeneous systems. 
Two most notable challenges for web service composition are ensuring interoperability of services and achieving Quality of Service (QoS) optimisation \cite{fensel2011semantic}. \textit{Interoperability} of web services presents challenge in syntactic and semantic dimensions. The syntactic dimension is covered by the XML-based technologies, such as $WSDL$, $SOAP$. The semantic dimension enables a better collaboration through ontology-based semantics, such as OWL-S, WSML, and SAWSDL \cite{petrie2016web}. \textit{Semantic web services composition} is distinguished from the syntactic service composition, as the resources of semantic web services are described semantically to enable a better interoperability for chaining web services. Another challenge is related to QoS optimisation. This problem gives birth to \textit{QoS-aware service composition} that aims to find composition solutions with optimised QoS.

Existing works on service composition focus mainly on addressing only one challenge above. In these works, huge efforts have been devoted to QoS-aware web service compositions assuming a pre-defined abstract workflow is given. This is generally considered as a \textit{semi-automated web service composition} approach. Generating composition plans automatically in discovering and selecting suitable web services is a NP-hard problem \cite{moghaddam2014service}. In the past few years, many approaches \cite{gupta2015optimization,ma2015hybrid,qi2010combining,da2016particle,da2015graphevol,yu2013adaptive} to QoS-aware web service composition employ Evolutionary Computation (EC) techniques to automatically generate composition solutions. Genetic Programming (GP) based approaches produce promising results, but these approaches often require repairing or penalising the solutions \cite{ma2015hybrid,yu2013adaptive}. However, Particle Swarm Optimisation (PSO) is considered to be more efficient, as it simply updates particles' positions to move candidate particles across the searching space, rather than employing genetic operators to carry out searching process. All these works have enabled an \textit{automatic web service composition}, but do not optimise QoS and quality of semantic matchmaking simultaneously to achieve a desirable balance on both sides.

The overall goal of this paper is to \textit{develop a PSO-based approach to comprehensive quality-aware automated semantic web service composition that satisfactorily optimises both QoS and semantic matchmaking quality}. Particularly, this paper extends existing works on QoS-aware service composition by considering jointly optimising the both quality aspects, which is proposed as a comprehensive quality model. We will achieve three objectives in this work:

@@

 \item To propose a PSO-based service composition approach using the proposed comprehensive quality model. To do that, we aim to find a service candidate queue that can be decoded into a service composition with near-optimal comprehensive quality.  
 \item To address the effectiveness of our PSO-based approach and a desirable balance achieved using our comprehensive quality model, we first compare our PSO-based approach with one recent GP-based approach \cite{ma2015hybrid} using our proposed quality model, and then compare our proposed quality model with one widely used QoS model using our proposed PSO-based approach.
\end{enumerate}
\vspace{-0.5cm}
\section{Related Work} \label{relatedWork}
While web service composition has attracted much research over the last decade, most efforts focus on either semantic web service composition \cite{bansal2016generalized,boustil2014semantic,mier2015integrated} or QoS-aware web service composition \cite{gupta2015optimization,ma2015hybrid,qi2010combining,da2016particle,da2015graphevol,yu2013adaptive}. Only very few works address both semantic matchmaking quality and QoS for web service composition problems. To the best of our knowledge, \cite{fanjiang2014semantic,lecue2009optimizing,pop2009immune} reported about first attempts that consider both aspects together.

@@

In summary, despite a large number of approaches for semantic web service composition and QoS-aware service composition approaches, there is a lack of a fully automated semantic service composition approach to optimise semantic matchmaking quality and QoS simultaneously. 
\vspace{-0.3cm}

\section{Problem Description and Comprehensive Quality Model}\label{problemDes}

@@

Given a service request $T=(I_T,O_T)$, we represent a service composition solution for $T$ with services $S_1,\ldots,S_n$  by a weighted DAG, $WG=(V,E)$ with node set $V=\{Start, S_1, S_2, \ldots, S_n, End\}$ and edge set $E = \{e_{1}, e_{2},... e_{m} \}$. $Start$ and $End$ are two special services defined as $Start = (\emptyset, I_T, \emptyset )$ and $End  = (O_T, \emptyset, \emptyset)$ that account for the input and output requirements given by the request. Each edge $e$ from a service $S$ to a service $S'$ means that service $S$ produces an output $a\in O_S$ that is matched ($exact$ or $plugin$) to an input $b\in I_{S'}$ to be consumed by service $S'$ in the composition. Based on the matchmaking type ($type_e$), the \emph{semantic similarity} ($sim_e$) of edge $e$ can be defined as follows:
\vspace{-0.2cm}
\begin{equation}
\label{equation10}
type_e = 
\begin{cases}
	1 & \text{ if $a\equiv b$ ($exact$ match)},\\
	p & \text{ if $a \sqsubseteq b$ ($plugin$ match)}
\end{cases}
\end{equation}
\begin{equation}
\label{equation11}
sim_e = sim(a,b) = \frac{2N_c}{N_{a}+N_{b}}
\end{equation}

\noindent with a suitable parameter $p, 0<p< 1$. $p$ is determined by user's preference, see Section \ref{experiment_design}. However, if more than one pair of matched output and input exist from service $S$ to service $S'$, $type_e$ and $sim_e$ will take on their average values.

The \emph{semantic matchmaking quality} of the service composition can be obtained by aggregating $type_e$ and $sim_e$, respectively, over all edges in $E$ as follows:
\vspace{-0.2cm}
\begin{equation}
\label{equation6}
MT {=} \prod_{j=1}^{m} type_ {e_{j}}
\end{equation}
\begin{equation}
\label{equation7}
SIM {=} \frac{1}{m}\sum_{j=1}^m sim_ {e_{j}}
\end{equation}
%\begin{equation}
%\label{eqCQ}
%CQ \stackrel{.}{=} (MT, \  SIM, \  A,\  R,\  C,\  T)
%\end{equation}
The \emph{QoS} of the service composition can be obtained by aggregating the QoS values of the participating services \cite{ma2015hybrid}. For a service composition with services $ S_1, S_2, ... S_n$ we obtain the reliability $R=\prod\limits^n_{k=1}r_{S_k}$, the availability $A=\prod\limits^n_{k=1}a_{S_k}$, the cost $C=\sum\limits^n_{k=1}c_{S_k}$, and the response time $T$ is the time of most time-consuming path in the composition, i.e., $$T = MAX \{\sum\limits^{\ell_j}_{k=1}t_{S_k} | \text{ $j\in\{1,\ldots,m\}$ and $\ell_j$ is the length of a path $P_j$}\}.$$

%Consequently, in Eq. \ref{eqCQ}, the comprehensive quality $CQ$ consists of the aggregated semantic matchmaking quality $SM$ and the QoS calculation in DAG discussed before.
\vspace{-0.1cm}
\noindent When multiple quality criteria are involved into decision making, then the overall fitness of a solution can be defined as a weighted sum of the individual criteria in Eq \ref{equation8}, assuming the preference of each quality criterion is known by users. 
\vspace{-0.2cm}
\begin{equation}
\label{equation8}
Fitness = w_1 \hat{MT} + w_2 \hat{SIM} + w_3 \hat{A} + w_4 \hat{R} + w_5(1 - \hat{T}) + w_6(1 - \hat{C})
\end{equation}
\noindent with $\sum_{k=1}^{6} w_k= 1$. We call this objective function the \emph{comprehensive quality model} for service composition.
The weights can be adjusted according to users' preferences. $\hat{MT}$, $\hat{SIM}$, $\hat{A}$, $\hat{R}$, $\hat{T}$, and $\hat{C}$ are normalised values calculated within the range from 0 to 1 using Eq. (\ref{equation9}). To simplify the presentation we also use the notation $(Q_1,Q_2,Q_3,Q_4,Q_5,Q_6) = (MT,SIM,A,R,T,C)$. $Q_1$ and $Q_2$ have minimum value 0 and maximum value 1. The minimum and maximum value of $Q_3$, $Q_4$, $Q_5$, and $Q_6$ are calculated across all task-related candidates in the service repository $\mathcal{SR}$ using the greedy search in \cite{ma2015hybrid,da2016genetic}.

@@

\section{PSO-based Approach to Comprehensive Quality-Aware Automated Semantic Web Service Composition}\label{qswsc_approach}
\subsection{An Overview of our PSO-based Approach}\label{PSO_based_approach}
As PSO has shown promise in solving combinatorial optimisation problems, we propose a PSO-based approach to comprehensive quality-aware automated semantic web service composition. Fig. \ref{overview} shows an overview of our approach consisting of four steps: 
\begin{figure}[h]
\centering
\fbox{\includegraphics[scale=.4]{overview.pdf}}
 \caption{An overview of our PSO-based approach to comprehensive quality-aware automated semantic web service composition.}
 \label{overview}
\end{figure}
\vspace{-0.5cm}

@@

Recently, a QoS Model, $Fitness = w_1 \hat{A} + w_2 \hat{R} + w_3(1 - \hat{T}) + w_4(1 - \hat{C})$, where $\sum_{i=1}^{4} w_i = 1$, is widely used for QoS-aware web service composition \cite{ma2015hybrid,da2016particle,da2015graphevol}. To show the effectiveness of our proposed comprehensive quality model, we compare the best solutions found by this QoS model and our comprehensive model using our PSO-based approach. We record and compare the mean values of both $SM$ ($SM = 0.5 \hat{MT} + 0.5 \hat{SIM}$) and $QoS$($QoS = 0.25 \hat{A} + 0.25 \hat{R} + 0.25(1 - \hat{T}) + 0.25(1 - \hat{C})$) of best solutions over 30 independent runs. To make the comparison informative, all these recorded values have been normalised from 0 to 1, and compared using independent-samples T tests, see Table \ref{decisionTable}. Note that p-values are smaller than 0.001 in the significance of the results.
